How cryptic is cryptic diversity? \newline Machine learning approaches to fine scale variation in the morphology of \textit{Emys marmorata}.

### 2013-06-12 documents/turtle_presentation.Rnw
presentation very close to finished

\documentclass{beamer}
\usepackage{amsmath, amsthm}
\usepackage{parskip, microtype, graphicx, caption, multirow}

@@

@@

\begin{frame}
  \frametitle{Morphological hypotheses}
\end{frame}

@@

  % Number of subspecies and where they occur.
  % images of the different phylobiogeographic hypotheses from the different spinks papers

@@

  (dissimilarity based) Evidence accumulation clustering
\end{frame}

@@

  Fancy way of saying classification and regression.

@@

  Multinomial logistic regression

  Random forests
\end{frame}

@@

\begin{frame}
  \frametitle{Second best cluster}
<<>>=
gap.second <- pam(as.dist(turtle.adult.dist), k = 2)
gap.map <- map.plot(data = turtle.adult,
                    label = gap.second$clustering,
                    map = map)
gap.map <- gap.map + scale_fill_manual(values = cbp)
gap.map <- gap.map + theme(legend.position = 'none')
gap.map
@

@@

\begin{frame}
  \frametitle{Generalize}
<<>>=
oo <- Reduce(cbind, Map(function(x) x$t, rr))
colnames(oo) <- groups
dd <- melt(oo)
gdist <- ggplot(dd, aes(x = value, fill = X2)) + geom_density(alpha = 0.3)
gdist <- gdist + theme(legend.title = element_blank())
gdist <- gdist + scale_fill_manual(values = cbp)

@@

\begin{frame}
  \frametitle{Best classification scheme?}
<<>>=
gg <- class.map(adult.train$spinks, adult.test$spinks, labs$spinks, 
                trf.a.analysis$class$spinks$pred, map)
gg <- gg + scale_fill_manual(values = cbp)
gg
@

@@

\begin{frame}[fragile]
  \frametitle{Variable importance}
  % ggpairs from the best random forest model
<<imp, fig.width = 3.5, fig.height = 3.5, eval = TRUE>>=
most.imp <- trf.a$spinks$optVariables
ww <- 'spinks'
ggimp <- ggpairs(turtle.adult,
                 columns = c(most.imp[1:3], ww), 
                 colour = 'spinks',
                 upper = 'blank',
                 params = c(LabelSize = 2, gridLabelSize = 2, size = 1))
ggimp
@
\end{frame}

\section{Discussion}
\begin{frame}
  \frametitle{Future}
\end{frame}

>>>>

\documentclass{beamer}
\usepackage{amsmath, amsthm}
\usepackage{parskip, microtype, graphicx, caption, multirow, subcaption}

@@

require(shapes)
require(geomorph)

@@

\begin{frame}
  \frametitle{Morphological hypothesis}

  \begin{figure}[h]
    \centering
    \captionsetup{justification = raggedleft, slc = off}
    \includegraphics[height = 0.8\textheight, keepaspectratio = true]{figure/seeliger}
    \caption*{\scriptsize{Seeliger 1945 \textit{Copeia}}}
    \label{fig:seeliger}
  \end{figure}
  % basic north/south stuff
  % earlier work by the masters student

@@

  \begin{figure}[h]
    \centering
    \captionsetup{justification = raggedleft, slc = off}
    \includegraphics[height = 0.8\textheight, keepaspectratio = true]{figure/spinks10}
    \caption*{\scriptsize{Spinks \textit{et al.} 2010 \textit{Molec. Ecol}}}
    \label{fig:spinks}
  \end{figure}
  % Number of subspecies and where they occur.
  % images of the different phylobiogeographic hypotheses from the different spinks papers

@@

  %(dissimilarity based) Evidence accumulation clustering

@@

  Fancy way of saying classification (and regression).

@@

  Multinomial logistic regression and random forests.

@@

\begin{frame}
  \frametitle{Second best cluster}
<<>>=
gap.second <- pam(as.dist(turtle.adult.dist), k = 2)
gap.map <- map.plot(data = turtle.adult,
                    label = gap.second$clustering,
                    map = map)
gap.map <- gap.map + scale_colour_manual(values = cbp)
gap.map <- gap.map + theme(legend.position = 'none',
                           legend.text = element_text(size = 5),
                           axis.title = element_text(size = 8),
                           axis.text = element_text(size = 5))
gap.map
@

@@

\begin{frame}
  \frametitle{Generalize}
<<>>=
oo <- Reduce(cbind, Map(function(x) x$t, rr))
colnames(oo) <- groups
dd <- melt(oo)
gdist <- ggplot(dd, aes(x = value, fill = X2)) + geom_density(alpha = 0.3)
gdist <- gdist + theme(legend.title = element_blank(),
                       legend.margin = unit(0, 'cm'),
                       legend.text = element_text(size = 5),
                       axis.title = element_text(size = 8),
                       axis.text = element_text(size = 5))
gdist <- gdist + scale_fill_manual(name = '', values = cbp)

@@

\begin{frame}
  \frametitle{Best classification scheme?}
<<>>=
gg <- class.map(adult.train$spinks, adult.test$spinks, labs$spinks, 
                trf.a.analysis$class$spinks$pred, map)
gg <- gg + scale_colour_manual(name = '', values = cbp)
gg <- gg + theme(#legend.position = 'none',
                 legend.margin = unit(0, 'cm'),
                 legend.text = element_text(size = 5),
                 axis.title = element_text(size = 8),
                 axis.text = element_text(size = 5))
gg
@

@@

\begin{frame}[fragile]
  \frametitle{Variable importance}
  % ggpairs from the best random forest model
<<imp, fig.width = 3.5, fig.height = 3.5, eval = TRUE>>=
most.imp <- trf.a$spinks$optVariables
ww <- 'spinks'
ggimp <- ggpairs(turtle.adult,
                 columns = c(most.imp[1:3], ww), 
                 colour = 'spinks',
                 upper = 'blank',
                 params = c(LabelSize = 2, gridLabelSize = 2, size = 1))

#putPlot
ggimp
@
\end{frame}

\begin{frame}
  \frametitle{Shape across most important features}

<<tps, include = FALSE>>=
fst.max <- which.max(turtle.adult[, most.imp[1]])
fst.min <- which.min(turtle.adult[, most.imp[1]])

snd.max <- which.max(turtle.adult[, most.imp[2]])
snd.min <- which.min(turtle.adult[, most.imp[2]])

par(mar = c(0, 0, 1, 1) + 0.1, oma = c(5, 4, 0, 0) + 0.1)

tpsgrid(turtle.land.adult[, , fst.max], mt)
tpsgrid(turtle.land.adult[, , fst.min], mt)

tpsgrid(turtle.land.adult[, , snd.max], mt)
tpsgrid(turtle.land.adult[, , snd.min], mt)

imp <- cbind(turtle.adult[, most.imp[1:2]], sp = turtle.adult$spinks)
imp.pca <- ggplot(imp, aes(x = PC3, y = PC8, colour = sp)) + geom_point()
imp.pca <- imp.pca + scale_colour_manual(name = '', values = cbp)
imp.pca
@

% show tps grids to the side of the PCA plot of the first two most important axes
\begin{figure}[h]
  \centering
  \hspace*{\fill}
  \begin{subfigure}[h]{0.3\textwidth}
    \centering
    \includegraphics[width = \textwidth]{figure/tps2}    
  \end{subfigure}
  \begin{subfigure}[h]{0.3\textwidth}
    \centering
    \includegraphics[width = \textwidth]{figure/tps5}    
  \end{subfigure}
  \begin{subfigure}[h]{0.3\textwidth}
    \centering
    \includegraphics[width = \textwidth]{figure/tps1}    
  \end{subfigure}
  \hspace*{\fill}
\end{figure}

\begin{frame}[fragile]
  \frametitle{Mean shape of classes}

<<fig.width = 3.75, fig.height = 3.75>>=
mt <- mshape(turtle.land.adult)
# get mean of each class
spi <- turtle.adult$spinks
wspi <- lapply(levels(spi), function(x, y) which(y == x), y = spi)
mspi <- lapply(wspi, function(x, y) mshape(y[, , x]), y = turtle.land.adult)
par(mfrow = c(2, 2), mar = c(0, 0, 1, 1) + 0.1, oma = c(5, 4, 0, 0) + 0.1,
    xaxt = 'n', yaxt = 'n')
mapply(function(x, m, n) {
      tpsgrid(x, m)
      title(main = paste(n))}, x = mspi, n = levels(spi),
      MoreArgs = list(m = mt), SIMPLIFY = FALSE)
par(mfrow = c(1, 1))
@

\section{Discussion}
\begin{frame}
  \frametitle{Future}

  \begin{itemize}
    \item illustration of morphological validation of previously cryptic variation
      \begin{itemize}
        \item the concordance is remarkable
        \item large sample sizes can be difficult
      \end{itemize}
    \item utility of large data, machine learning methods 
    \item unsupervised methods for when no explicit hypothesis -- nonparametric Bayes
    \item cause of interclass variation -- local adaptation? pure isolation?
  \end{itemize}
